detect patches

### 2024-02-06 scripts/classify_patches.Rmd
fn to note occupied patches

```{r}
library(terra)
library(landscapemetrics)
```

@@

@@

lapply(f[21:335], FUN = landscapR, pout = 'patch_metrics', pRout = 'fine_patches')

@@

f <- file.path('../results/fine_patches', list.files('../results/fine_patches'))

recs <- st_read('../data/raw/occurrence/combined_records') %>% 
  filter(PtType == 1)

@@

  r <- terra::rast(f)
  r_taxon <- gsub('[.]tif', '', basename(f))
  occs <- dplyr::filter(recs, taxon == r_taxon) |>
    dplyr::mutate(ID = 1:nrow(.)) |>
    terra::vect()

@@

  # if pts were NA, buffer to 90m, many occupied areas were 'cut' away by the 
  # hydrologic mapping. 
  occ_poly <- dplyr::filter(occ_pts, is.na(RASTER)) |>
    sf::st_buffer(occ_pts, 120) # we cut out 90, but let's go a bit further for corner cases.
  occ_poly <- terra::extract(r, occ_poly, bind = TRUE, weights = TRUE) |>
    sf::st_as_sf() |>
    sf::st_centroid()  |>
    dplyr::group_by(ID) |> # now make sure we get the patch with the 'most' overlap with the polygon
    dplyr::slice_max(WEIGHT) |>
    dplyr::drop_na(RASTER) # these values actually NA.

  occ_pops <- dplyr::bind_rows( # a dataframe of all raster patch ID's with populations known from within
    # the last few decades in them. 
    dplyr::drop_na(occ_pts), occ_poly) |>
    dplyr::distinct(RASTER, .keep_all = TRUE) # this is multiple pts per patch - not neccessary for us.

  write.csv(occ_pops, row.names = F,
            file.path('../results/occupiedPatches', paste0(r_taxon, '.csv')))

@@

>>>>

```{r}
library(terra)
library(tidyverse)
library(sf)
library(landscapemetrics)
```

@@

Create finer scale patches which more or less are meant to represent populations

```{r eval = F}

@@

lapply(f, FUN = landscapR, pout = 'patch_metrics', pRout = 'fine_patches')

@@

f <- file.path('../results/fine_patches', list.files('../results/fine_patches'))
p <- '../data/raw/occurrence/combined_records'

recs <- lapply(
  file.path(p, list.files(p, pattern = '.shp$')), sf::st_read, quiet = TRUE) %>% 
  bind_rows() %>% 
  filter(Occurrence == 1)

r1 <- rast(f[1])

@@

  r <- terra::rast(x)
  r_taxon <- gsub('_', ' ', gsub('-patchIDS[.]tif', '', basename(x)))
  occs <- dplyr::filter(recs, taxon == r_taxon) |>
    dplyr::mutate(ID = dplyr::row_number()) |>
    terra::vect()

@@

  # if pts were NA, buffer to 90m, many occupied areas were 'cut' away by the 
  # hydrologic mapping. 
  occ_poly <- dplyr::filter(occ_pts, is.na(lyr.1)) |>
    sf::st_buffer(250) # we cut out 90, but let's go a bit further for corner cases.

  occ_poly <- terra::extract(r, occ_poly, weights = TRUE) %>% 
    dplyr::group_by(ID) |> # now make sure we get the patch with the 'most' overlap with the polygon
    tidyr::drop_na(lyr.1) |> # these values actually NA. 
    dplyr::slice_max(weight, with_ties = F) |>
    dplyr::left_join(filter(occ_pts) |>
                       select(-lyr.1), by = join_by('ID')) |>
    dplyr::select(-weight)

  occ_pops <- dplyr::bind_rows( # a data frame of all raster patch ID's with populations known from within
    # the last few decades in them. 
    tidyr::drop_na(occ_pts, lyr.1), occ_poly) |>
    dplyr::rename(PatchID = lyr.1)

  # write out summary values of the patch matching
  data.frame(
    noPres = , 
    PresTopSuitability =, # the number of presences located in top suitability patches
    UniquePatchID = , # the number of uniquelly identified patches
    NoPatchesTopSuitability =  # the total number of most suitable patches from the RASTER
  )

  occ_pops <- dplyr::distinct(occ_pops, PatchID, .keep_all = TRUE) |> # this is multiple pts per patch - not necessary for us. 
    sf::st_drop_geometry() |>
    dplyr::select(date, PatchID) |>
    dplyr::arrange(PatchID) |>
    write.csv(row.names = F,
            file.path('../results/occupiedPatches-topSuitability', paste0(r_taxon, '.csv')))

  # for points which were not patched to a most suitable patch, put them into a hydrologic basin

@@

patchTaggR(f[1])
```

```{r}
PLANTS <- read.csv('../data/raw/2024sos_species.csv') |>
  select(genus:symbol) |>
  unite(sctfcnm, genus:infraspecies, na.rm = TRUE)
